A crop modelling strategy to improve cacao quality and productivity

### 2021-12-15 paperModel/paperCacaoModel_round1.tex
round1 corrections

\abstract{Cacao production systems in Latin America have a high importance over social and economic development, facing  the fight against hunger and poverty. Although Colombian cacao has the potential to be in the high value markets for fine flavour, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. \textcolor{blue}{Traditionally cacao harvest takes place at 5 or 6 months  by calendar after flowering, the problem is that  the seeds are often  unripe or  over ripe germinating inside the pod afecting the quality of the final cocoa products}. Other environmental parameters that have more association with pod maturation speed are not taken into account.  Cacao fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments. In this context, crop models are useful tools to simulate and predict crop development over time and under multiple environmental conditions. Since, harvesting at the right time can yield high quality cacao, we parametrised a crop model to predict the best time for harvest cocoa fruits in Colombia. . Results can be a practical tool that supports cacao farmers in the production of high quality cocoa.  When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions. This crop model contributed to understand and predict the phenology of cacao fruits for two key cultivars ICS95 y CCN51. \textcolor{blue}{The aim of this study was to calibrate a crop model that simulates cacao crop development and yield, and predicts the maturation day when the fruit is likely to be ready for harvest.}}

@@

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. The farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented. \textcolor{blue}{The problem is that  the seeds are often  unripe or  over ripe germinating inside the pod affecting the cocoa quality}. This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the final cocoa product \citep{Escobar2021}.  \textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.  Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits.}

>>>>

\abstract{Cacao production systems in Latin America have a high importance over social and economic development, facing  the fight against hunger and poverty. Although Colombian cacao has the potential to be in the high value markets for fine flavour, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. \textcolor{blue}{Traditionally cacao harvest takes place at 5 or 6 months  by calendar after flowering, the problem is that  the seeds are often  unripe or  over ripe germinating inside the pod afecting the quality of the final cocoa products}. Other environmental parameters that have more association with pod maturation speed are not taken into account.  Cacao fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments. In this context, crop models are useful tools to simulate and predict crop development over time and under multiple environmental conditions. Since, harvesting at the right time can yield high quality cacao, we parametrised a crop model to predict the best time for harvest cocoa fruits in Colombia. . Results can be a practical tool that supports cacao farmers in the production of high quality cocoa.  When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions. This crop model contributed to understand and predict the phenology of cacao fruits for two key cultivars ICS95 y CCN51. \textcolor{blue}{The aim of this study was to calibrate a crop model that simulates cacao crop development and yield, and predicts the maturation day when the fruit is likely to be ready for harvest which will improve the quality of the seeds.}}

@@

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. The farmer practice of empirically harvesting from 5 to 6 months or 180 Days After Flowering  (DAF) date, produced a mix of quality of cacao beans that are then fermented. \textcolor{blue}{The problem is that  the seeds are often  unripe or  over ripe germinating inside the pod affecting the cocoa quality}. This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the final cocoa product \citep{Escobar2021}.  \textcolor{blue}{Therefore,  It is important to identify the best moment to harvest  considering physiological responses affected by  soil types , cultivar specifications and weather variables such as rain, solar radiation and wind.  Thus,  for cacao production in Colombia, physiological simulation models may be valuable to predict the right moment to harvest cocoa fruits, hence it will improve the quality of seeds.}
